---
output: github_document #rmarkdown::html_vignette
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# text

<!-- badges: start -->
<!-- badges: end -->

The language that individuals use contains a wealth of information interesting for research. *Text* analyzes and visualizes text, and their relation to other text or numerical variables. The r-package is based on state-of-the-art techniques from statistics and artificial intelligence, including natural language processing, deep learning and machine learning. 

*Text* is created through a collaboration between psychology and computer science to address research needs and ensure state-of-the-art techiniques. It provides powerful functions tailored to test research hypotheses in social and behaviour sciences for both relatively small and large datasets.

### Short installation guide
Most users simply need to run below installation code. 
For those experiencing problems, please see the [Extended Installation Guide](https://www.r-text.org/articles/Extended%20Installation%20Guide.html).

[CRAN](https://CRAN.R-project.org) version:

``` r
install.packages("text")
```

[GitHub](https://github.com/) development version:

``` r
# install.packages("devtools")
devtools::install_github("oscarkjell/text")
```

### Map your text to numbers
Recent significant advances in NLP research have resulted in improved representations of human language (i.e., language models). These language models have produced big performance gains in tasks related to understanding human language. Using deep learning to model word order and context now gives great results. Multilingual language models can also represent several languages; multilingual BERT comprises *104 different languages*. Text are making these SOTA models  easily accessible through an interface to [HuggingFace](https://huggingface.co/transformers/) in Python.

*Table 1. Some of the available language models*
``` {r HF_tabble_short, echo=FALSE, results='asis'}
library(magrittr)
model_short <- c("BertModel", "BertModel", "XLNetModel", "DistilBertModel", "RobertaModel")

tokenizer_short <- c("BertTokenizer", "BertTokenizer", "XLNetTokenizer", "DistilBertTokenizer", "RobertaTokenizer")

pretrained_weights_shortcut_short <- c("'bert-base-uncased'","**'bert-multilingual-uncased'**","'xlnet-base-cased'",
                                 "'distilbert-base-cased'", "'roberta-base'")

HF_tables_short <- tibble::tibble(model_short, tokenizer_short, pretrained_weights_shortcut_short)

knitr::kable(HF_tables_short, caption="", bootstrap_options = c("hover"), full_width = T)
```
  
See [HuggingFace's Github](https://github.com/huggingface/transformers) for a more comprehensive list of models. 

### An end-to-end package
Text also provides functions to analyse the word embeddings, with well-tested machine learning algorithms and statistics. An example is how it is possible to plot statistically significant words in the word embedding space. 

```{r DPP_plot, message=FALSE, warning=FALSE}
library(text)
# Use data (DP_projections_HILS_SWLS_100) that have been pre-processed with the textProjectionData function

# The test-data included in the package is called: sq_data_tutorial_plotting_hilswl
plot_projection <- textProjectionPlot(
  word_data = DP_projections_HILS_SWLS_100,
  k_n_words_two_test = TRUE, 
  plot_n_words_square = 3,
  plot_n_words_p = 3,
  plot_n_word_extreme = 1,
  plot_n_word_frequency = 1,
  plot_n_words_middle = 1,
  x_axes = "dot.x",
  y_axes = "dot.y",
  p_values_x = "p_values_dot.x",
  p_values_y = "p_values_dot.y",
  p_alpha = 0.05,
  title_top = " Dot Product Projection (DPP) of Harmony in life words",
  x_axes_label = "Low vs. High HILS score",
  y_axes_label = "Low vs. High SWLS score",
  p_adjust_method = "bonferroni",
  scale_x_axes_lim = c(-15, 15),
  scale_y_axes_lim = c(-15, 15),
  y_axes_values_hide = FALSE
)
plot_projection

```


### References
Selected articles analysing human language using (decontextualized) word embeddings.

***Methods Articles***  
[Gaining insights from social media language: Methodologies and challenges](http://www.peggykern.org/uploads/5/6/6/7/56678211/kern_2016_-_gaining_insights_from_social_media_language-_methodologies_and_challenges.pdf).  
*Kern et al., (2016). Psychological Methods.*

[Semantic measures: Using natural language processing to measure, differentiate, and describe psychological constructs.](https://www.ncbi.nlm.nih.gov/pubmed/29963879) [Pre-print](https://psyarxiv.com/er6t7/)  
*Kjell et al.,  (2019). Psychological Methods.*

***Clinical Psychology***  
[Facebook language predicts depression in medical records](https://www.pnas.org/content/115/44/11203)  
*Eichstaedt, J. C., ... & Schwartz, H. A. (2018). PNAS.*


***Social and Personality Psychology***     
[Personality, gender, and age in the language of social media: The open-vocabulary approach](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0073791)  
*Schwartz, H. A., … & Seligman, M. E. (2013). PloSOne.*

[Automatic Personality Assessment Through Social Media Language](https://www.semanticscholar.org/paper/Automatic-personality-assessment-through-social-Park-Schwartz/280034e01f41d0184b977479a5474bd805c2922c)  
*Park, G., Schwartz, H. A., ... & Seligman, M. E. P. (2014). Journal of Personality and Social Psychology.*


***Health Psychology***  
[Psychological language on Twitter predicts county-level heart disease mortality](https://journals.sagepub.com/doi/abs/10.1177/0956797614557867)  
*Eichstaedt, J. C., Schwartz, et al. (2015). Psychological Science.*


***Positive Psychology***  
[The Harmony in Life Scale Complements the Satisfaction with Life Scale: Expanding the Conceptualization of the Cognitive Component of Subjective Well-Being](https://link.springer.com/article/10.1007/s11205-015-0903-z)  
*Kjell, et al., (2016). Social Indicators Research*

***Computer Science: Python Software***  
[DLATK: Differential language analysis toolkit](http://aclweb.org/anthology/D17-2010)  
*Schwartz, H. A., Giorgi, et al., (2017).  In Proceedings of the 2017 Conference on Empirical Methods in Natural Language Processing: System Demonstrations* 


[DLATK](http://dlatk.wwbp.org/)



