---
title: "Extended Installation Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extended_Installation_Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*Text* uses an r-package called [reticulate](https://rstudio.github.io/reticulate/) as an interface to [Python](https://www.python.org/), and the Python packages [torch](https://pytorch.org/) and [transformers](https://huggingface.co/transformers/). Reticulate normally installs these necessary packages automatically, however, if you receive errors indicating either of these packages are missing follow the instructions below.

```{r extended_installation, eval = FALSE}
##library(text)
##
### To install the python packages torch, transformers, numpy and nltk through R, run: 
##library(reticulate)
##install_miniconda()
##
##conda_install(envname = 'r-reticulate', c('torch', 'transformers', 'numpy', 'nltk'), pip = TRUE)
##
### Windows 10
##conda_install(envname = 'r-reticulate', c('torch', 'transformers', 'numpy', 'nltk'))

```


##Checking your versions
If something isn't working right, it is a good start to examine what is installed and running on your system. For example to make sure that you have R and Python versions that are up to date. 

```{r extended_installation_checking_system, eval = FALSE}
##
### First check R-version and which packages that are attached and loaded.  
##sessionInfo()
##
### Second check out python version; and make sure you at least have version 3.6.10
##library(reticulate)
##py_config()
```


## Issue: RStudio craches during textEmbed
After a new install/update of *text*, RStudio crashed (Abort session) when running functions that fetches word embeddings (i.e., ```textEmbedLayersOutput``` or ```textEmbed```). 

### Solution: Reinstall reticulate and r-miniconda
To solve the issue re-install reticulate (development version) and uninstall and install r-miniconda.  

Uninstall r-miniconda by removing its entire folder (which by default [in Mac] is at ```Users/YOUR_USER_NAME/Library/r-miniconda```). 

(Note that [in Mac] the Library folder is hidden, so to make it visible go to Finder and the path Users/YOUR_USER_NAME/ and press the three keys: ```COMMAND + SHIFT + .``` . Then the Library-folder should appear, and you can find and remove r-miniconda.


```{r extended_REinstallation, eval = FALSE}
##library(text)
##
### To re-install packages start with a fresh session by restarting R and RStudio
##
### Install development of reticulate (might not be necessary)
##devtools::install_github("rstudio/reticulate")
##
### After having manually removed the r-miniconda folder, install it again: 
##library(reticulate)
##install_miniconda()
##
### Subsequently re-install torch, transformers, numpy and nltk by running: 
##conda_install(envname = 'r-reticulate', c('torch', 'transformers', 'numpy', 'nltk'), pip = TRUE)
##
```


The exact way to install these packages may differ across systems. Please see:   
[Python](https://www.python.org/)   
[torch](https://pytorch.org/)  
[transformers](https://huggingface.co/transformers/)  
[transformers installation](https://huggingface.co/transformers/installation.html)

## Share advise
*If you find a good solution please feel free to email [us](oscar.kjell@psy.lu.se) so that we can update above instructions.*
 
 
 





