---
title: "Extended Installation Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extended_Installation_Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Text uses an r-package called [reticulate](https://rstudio.github.io/reticulate/) as an interface to [Python](https://www.python.org/) and the Python packages [torch](https://pytorch.org/) and [transformers](https://huggingface.co/transformers/). Reticulate normally installs these necessary packages automatically, however, if you receive errors indicating either of these packages are missing see below.

```{r extended_installation, eval = FALSE}
library(text)

# To manually install the python packages torch and transformers through R, run: 
library(reticulate)
install_miniconda()
conda_install(envname = 'r-reticulate', c('torch', 'transformers==2.4.1'), pip = TRUE)

# Windows 10
conda_install(envname = 'r-reticulate', c('PyTorch', 'transformers'))

```

The exact way to install these packages may differ across systems. If the above does not work for you try installing the python packages outside of R or RStudio. Please see: 
[Python](https://www.python.org/) 
[torch](https://pytorch.org/)
transformers
[transformers](https://huggingface.co/transformers/) and [transformers installation](https://huggingface.co/transformers/installation.html)

If you find a good solution please feel free to email it to [Oscar](oscar.kjell@psy.lu.se) for inclusion above. 



