---
title: "Pre-Trained Models"
vignette: >
  %\VignetteIndexEntry{Pre-Trained Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Search the table below for pre-trained models that can be used with the `textPredict()` function. The models have been trained using one of the `textTrain()` functions and can now be used with `textPredict()` by providing the `model_info` parameter with the URL from the `Path` column in the table below. In some cases you can also provide the `model_info` parameter with the name from the `Name` column (although, for this to work the name needs to be coupled with its URL within the code of the `textPredict()` function).

If you want to add a pre-trained model to the table, please fill out the details in this [Google sheet](https://docs.google.com/spreadsheets/d/1K16JdK7zOmuRktqgaYs5sgaHnkUVwB9v6ERxZdd9iW8/edit#gid=0) and contact us (email oscar [ d_o t] kjell [a _ t] psy [DOT] lu [d_o_t]se) so that we can update the website. 

Note that you can adjust the width of the columns when scrolling the table.

```{r models_table, eval = TRUE, echo=FALSE}
library("reactable")
# see vignette: https://glin.github.io/reactable/articles/examples.html#custom-rendering

model_data <- read.csv(system.file("extdata",
                                   "text_models_info.csv",
                                   package = "text"),
                       header = TRUE, 
                       skip = 2)

reactable::reactable(
  data = model_data,
  filterable = TRUE,
  defaultPageSize = 50,
  highlight = TRUE, 
  resizable = TRUE
)
```

