---
title: "Getting started"
output: github_document #rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{text}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The *text*-package uses natural language processing and machine learning methods to examine text and numerical variables.  

Central *text* functions are described below. The data and methods come from the [Kjell et al., 2019](https://www.ncbi.nlm.nih.gov/pubmed/29963879) ([pre-print](https://psyarxiv.com/er6t7/)), which show how individuals' open-ended text answers can be used to *measure*, *describe* and *differentiate* psychological constructs.  

In short the workflow includes to first transform text variables into word embeddings. These word embeddings are then used to, for example, predict numerical variables, compute semantic similarity scores, statistically test difference in meaning between two sets of texts and plot words in the word embedding space.  

### textEmbed(): mapping text to numbers
The ```textEmbed()``` function automatically transforms character variables in a given tibble to word embeddings. The example data that will be used in this tutorial comes from participants that have described their harmony in life and satisfaction with life with a text response, 10 descriptive words or rating scales. For a more detailed description please see the [word embedding tutorial](https://www.r-text.org/articles/Word%20embeddings.html) 

```{r setup, eval = FALSE, warning=FALSE, message=FALSE}

library(text)

# Get example data including both text and numerical variables
sq_data <- Language_based_assessment_data_8_10

# Transform the text data to BERT word embeddings
wordembeddings <- textEmbed(sq_data)

# See how word embeddings are structured
wordembeddings

# Save the word embeddings to avoid having to import the text every time
saveRDS(wordembeddings, "wordembeddings.rds")

# Get the word embeddings again
wordembeddings <- readRDS("_YOURPATH_/wordembeddings.rds")
```


### textTrain(): Examine the relationship between text and numeric variables
The ```textTrain()``` is used to examine how well the word embeddings from a text can predict a numeric variable. This is done by training the word embeddings using ridge regression and 10-fold cross-validation (where the word embeddings are pre-processed using pca). In the example below we examine how well the harmony text responses can predict the rating scale scores from the Harmony in life scale.

```{r, eval = FALSE,  warning=FALSE, message=FALSE}
library(text)
library(rio)
# Load data that has already gone through textEmbed
# The previous example only imported 10 participants; whereas below we load data from 100 participants
wordembeddings <- rio::import("https://r-text.org/text_data_examples/wordembeddings4_100.rda")
# Load corresponding numeric variables
numeric_data <-   rio::import("https://r-text.org/text_data_examples/Language_based_assessment_data_8_100.rda")

# Examine the relationship between harmonytext and the corresponding rating scale
model_htext_hils <- textTrain(wordembeddings$harmonytexts, numeric_data$hilstotal, nrFolds_k = 10)

# Examine the correlation between predicted and observed Harmony in life scale scores
model_htext_hils$correlation
```

## textDiff(): Test the difference in meaning between to sets of texts
The ```textDiff()``` function provides a permutation based test to examine whether two sets of texts significantly differ in meaning. It produces a p-value and estimate as an effect size. Below we examine whether the harmony text and satisfaction text responses differ in meaning. 
```{r, eval = TRUE, warning=FALSE, message=FALSE}
library(text)

# Compare the meaning between individuals' harmony in life and satisfaction with life answers
textDiff(wordembeddings4_10$harmonytexts, wordembeddings4_10$satisfactiontexts, Npermutations = 100, output.permutations = FALSE)
```

## Plot statistically significant words 
The plotting is made in two steps: First the ```textPlotData()``` function is pre-procssing the data, including computing statistics for each word to be plotted. Second, ```textPlotViz()``` is plotting the words, including many options to set color, font etc for the figure. Dividing this procedure into two steps makes the process more transparent (since the user naturally get to see the output that the words are plotted according to) and quicker since the more heavy computations are made in the first step, the last step goes quicker so that one can try different design settings.   

### textPlotData(): Pre-process data for plotting 
```{r, eval = FALSE, warning=FALSE, message=FALSE}
library(text)

# Pre-process word data to be plotted with textPlotViz-function
# Get example data within package
wordembeddings_10 <- wordembeddings4_10
data_10 <- sq_data_tutorial8_10

# Pre-process data
df_for_plotting <- textPlotData(data_10$harmonywords, wordembeddings_10$harmonywords,
  wordembeddings_10$singlewords_we,
  data_10$hilstotal, data_10$swlstotal
)
df_for_plotting
```


### textPlotViz(): A two-dimensional word plot
```{r, eval = TRUE, warning=FALSE, message=FALSE, dpi=300}
library(text)
# Use data (DP_projections_HILS_SWLS_100) that have been pre-processed with the textProjectionData function
plot_projection <- textProjectionPlot(
  word_data = DP_projections_HILS_SWLS_100,
  k_n_words_two_test = FALSE,
  plot_n_words_square = 5,
  plot_n_words_p = 5,
  plot_n_word_extreme = 1,
  plot_n_word_frequency = 1,
  plot_n_words_middle = 1,
  x_axes = "dot.x",
  y_axes = "dot.y",
  p_values_x = "p_values_dot.x",
  p_values_y = "p_values_dot.y",
  p_alpha = 0.05,
  title_top = " Dot Product Projection (DPP) of Harmony in life words",
  x_axes_label = "Low vs. High HILS score",
  y_axes_label = "Low vs. High SWLS score",
  p_adjust_method = "bonferroni",
  points_without_words_size = 0.4,
  points_without_words_alpha = 0.4,
  y_axes_values_hide = FALSE
)
plot_projection

```


### Relevant References
*Text* is new and has not been used in a publication yet. therefore, the below list consists of papers analysing human language in a similar fashion that is possible *text*.

***Methods Articles***  
[Gaining insights from social media language: Methodologies and challenges](http://www.peggykern.org/uploads/5/6/6/7/56678211/kern_2016_-_gaining_insights_from_social_media_language-_methodologies_and_challenges.pdf).  
*Kern et al., (2016). Psychological Methods.*

[Semantic measures: Using natural language processing to measure, differentiate, and describe psychological constructs.](https://www.ncbi.nlm.nih.gov/pubmed/29963879) [Pre-print](https://psyarxiv.com/er6t7/)  
*Kjell et al.,  (2019). Psychological Methods.*

***Clinical Psychology***  
[Facebook language predicts depression in medical records](https://www.pnas.org/content/115/44/11203)  
*Eichstaedt, J. C., ... & Schwartz, H. A. (2018). PNAS.*


***Social and Personality Psychology***     
[Personality, gender, and age in the language of social media: The open-vocabulary approach](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0073791)  
*Schwartz, H. A., … & Seligman, M. E. (2013). PloSOne.*

[Automatic Personality Assessment Through Social Media Language](https://www.semanticscholar.org/paper/Automatic-personality-assessment-through-social-Park-Schwartz/280034e01f41d0184b977479a5474bd805c2922c)  
*Park, G., Schwartz, H. A., ... & Seligman, M. E. P. (2014). Journal of Personality and Social Psychology.*


***Health Psychology***  
[Psychological language on Twitter predicts county-level heart disease mortality](https://journals.sagepub.com/doi/abs/10.1177/0956797614557867)  
*Eichstaedt, J. C., Schwartz, et al. (2015). Psychological Science.*


***Positive Psychology***  
[The Harmony in Life Scale Complements the Satisfaction with Life Scale: Expanding the Conceptualization of the Cognitive Component of Subjective Well-Being](https://link.springer.com/article/10.1007/s11205-015-0903-z)  
*Kjell, et al., (2016). Social Indicators Research*

***Computer Science: Python Software***  
[DLATK: Differential language analysis toolkit](http://aclweb.org/anthology/D17-2010)  
*Schwartz, H. A., Giorgi, et al., (2017).  In Proceedings of the 2017 Conference on Empirical Methods in Natural Language Processing: System Demonstrations* 


[DLATK](http://dlatk.wwbp.org/)



